---
title: "Assignment1"
author: "Manish Kumar"
date: "2023-01-11"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

### 1. Read Chapter 1 of the book

### 2. Download and analyze adjusted excess returns of S&P 500. In case if package `quantmod` is not working on your system, download the data directly from the web site <https://finance.yahoo.com>

#### 2.1 Calculate continuous daily excess returns of SP500 ("\^GSPC") for the period from 1/1/2022 until 12/22/2022 using overnight Fed Funds rates as risk-free rates

```{r}
# Load the required library
suppressMessages(library(quantmod))

# Get the S&P500 data and Fed Funds Rates
getSymbols("^GSPC", src="yahoo", from=as.Date("2022-01-01"), to=as.Date("2022-12-22"), periodicity="daily")
getSymbols("DFF", src="FRED", from=as.Date("2022-01-01"), to=as.Date("2022-12-22"), periodicity="daily")

# Calculate daily returns (dailyReturns automatically takes care of the Adjusted closing price)
SP500 <- dailyReturn(GSPC)

# Filter and order by the date range in S&P500 data
fed_funds_rate = DFF[rownames(data.frame(SP500))] / 100

# Calculate excess returns
daily_excess_returns = (SP500 - fed_funds_rate) * 100

# Show and plot the data
head(daily_excess_returns)
plot(daily_excess_returns, ylab="Excess returns of SP500", main="Continuous Excess Daily Returns of S&P 500")

```

#### 2.2 Calculate average daily excess return, actual return of S&P 500 in 2022 per day, and average FedFund rate of return per day

```{r}
# Average daily excess return
cat("Average Daily Excess Return:", mean(daily_excess_returns), "\n")

# Actual return for S&P500
head(SP500)
plot(SP500)

# Average FedFund rate of return
cat("Average FedFund rate of Return:", mean(fed_funds_rate), "\n")
```

### 3. Download and analyze exchange rates. In case if package `quantmod` is not working on your system, download the data directly from the web site <https://finance.yahoo.com>.

Answer the following questions (as in Exercise 5 on page 37) as a refresher of statistical analysis skills. Try to do it without using R demo code from the book.

#### 3.1 Download from `Oanda` using method `getFX()` from library `quantmode` for the period from July 1, 2022 to December 22, 2022:

-   GBP/USD exchange rate, i.e. price of 1 British pound in US dollars
-   USD/JPY exchange rate, i.e. price of 1 US dollar in Japanese yen

```{r}
# Download exchange rates
suppressWarnings(getFX("GBP/USD", src="oanda", from=as.Date("2022-07-01"), to=as.Date("2022-12-22")))
suppressWarnings(getFX("USD/JPY", src="oanda", from=as.Date("2022-07-01"), to=as.Date("2022-12-22")))
```

#### 3.2 Calculate daily log returns of both exchange rates

```{r}
daily_log_ret_gbp <- diff(log(GBPUSD))
daily_log_ret_jpy <- diff(log(USDJPY))

head(daily_log_ret_gbp)
head(daily_log_ret_jpy)
```

#### 3.3 Calculate sample min, mean, sd, skewness, kurtosis, max of log returns for both exchange rates

```{r}
suppressMessages(library(fBasics))

print("Stats for GBP")
# Min
cat("Min: ", min(daily_log_ret_gbp, na.rm=TRUE), "\n")

# Mean
cat("Mean: ", mean(daily_log_ret_gbp, na.rm=TRUE), "\n")

# Standard Deviation
cat("Sd: ", sd(daily_log_ret_gbp, na.rm=TRUE), "\n")

# Skewness
cat("Skewness: ", skewness(daily_log_ret_gbp, na.rm=TRUE), "\n")

# Kurtosis
cat("Kurtosis: ", kurtosis(daily_log_ret_gbp, na.rm=TRUE), "\n")

# Max
cat("Max: ", max(daily_log_ret_gbp, na.rm=TRUE), "\n")
```

```{r}
print("Stats for JPY")

# Min
cat("Min: ", min(daily_log_ret_jpy, na.rm=TRUE), "\n")

# Mean
cat("Min: ", mean(daily_log_ret_jpy, na.rm=TRUE), "\n")

# Standard Deviation
cat("Min: ", sd(daily_log_ret_jpy, na.rm=TRUE), "\n")

# Skewness
cat("Min: ", skewness(daily_log_ret_jpy, na.rm=TRUE), "\n")

# Kurtosis
cat("Min: ", kurtosis(daily_log_ret_jpy, na.rm=TRUE), "\n")

# Max
cat("Min: ", max(daily_log_ret_jpy, na.rm=TRUE), "\n")
```

#### 3.4 Test hypothesis H~0~: μ=0 against alternative H~0~: μ≠0

```{r}
suppressWarnings(t.test(daily_log_ret_gbp))
suppressWarnings(t.test(daily_log_ret_jpy))
```
In both cases, Since this p-value is greater than 0.05, we fail to reject the null hypothesis.
