---
title: "Assignment8"
author: "Manish Kumar"
date: "Feb 28, 2023"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

## Part I - Option Pricing
A stock index is currently at 810 and has volatility of 20%.

The risk-free rate is 5% per year.

Assume that dividend yield equals 1%.

#### 1. Price European 6-month call option with strike 800 using `EuropeanOption()` from `RQuantLib`

```{r}
suppressMessages(library(RQuantLib))

underlying = 810
volatility = 0.20
riskFreeRate = 0.05
dividendYield = 0.01
strike = 800
maturity = 0.5

call = EuropeanOption(
  type="call", underlying=underlying, strike=strike, riskFreeRate=riskFreeRate,
  dividendYield=dividendYield, maturity=maturity, volatility=volatility
)

call$value
```

#### 2. Calculate the same premium manually using the formulas on the last slide of the lecture notes.

Think how dividend yield should affect option price.

Experiment with the function `EuropeanOption()` with zero or non-zero dividend yield and find how the Black-Scholes formula on slide 17 should be modified for dividend yield.

$C=S_{0}N(d_{1})-(Kexp(-RT))N(d_{2})$

$d_{1}=(ln(S_{0}/K) + (R + .5\sigma^2)  T)/(\sigma \sqrt(T))$

$d2=d_{1} - (\sigma \sqrt(T))$

```{r}

d1 = (log(underlying / strike) + (riskFreeRate + .5 * (volatility^2)) * maturity) / (volatility * sqrt(maturity))
d2 = d1 - (volatility * sqrt(maturity))

C = (underlying * pnorm(d1, 0, 1)) - ((strike * exp(-riskFreeRate * maturity)) * pnorm(d2, 0, 1))
C
```


```{r}
# When dividendYield = 0
callDiv0 = EuropeanOption(
  type="call", underlying=underlying, strike=strike, dividendYield=0,
  riskFreeRate=riskFreeRate, maturity=maturity, volatility=volatility
)

callDiv0$value
```

As dividend yields increase, the call option price should decrease.

The Black Scholes Merton formula did not take dividend yields into consideration. The following is the Merton extension on the original formula.


$C=S_{0} exp(-qT)N(d_{1})-(Kexp(-RT))N(d_{2})$

$d_{1}=(ln(S_{0}/K) + (R - q+ .5\sigma^2)  T)/(\sigma \sqrt(T))$

$d2=d_{1} - (\sigma \sqrt(T))$

Where `q` is the dividend yield

```{r}


d1New = (log(underlying / strike) + (riskFreeRate - dividendYield + .5 * (volatility^2)) * maturity) / (volatility * sqrt(maturity))
d2New = d1New - (volatility * sqrt(maturity))

CNew = (underlying *exp(-dividendYield * maturity) * pnorm(d1New, 0, 1)) - ((strike * exp(-riskFreeRate * maturity)) * pnorm(d2New, 0, 1))

CNew
```

The new formula generated call option price identical to the answer in part 1. 

#### 3. Calculate the premium of put option with the same expiry and same strike using put-call parity

Put-Call parity without dividend yield:

$Put = Call + K exp(-RT) - S_{0}$

Put-Call parity with dividend yield:

$Put = Call + K exp(-RT) - S_{0}  exp(-qT)$

```{r}
putParity = CNew + (strike * exp(-riskFreeRate * maturity)) - (underlying * exp(-dividendYield * maturity))
putParity
```

Let's verify the Put-Call parity formula using `EuropeanOption()`

```{r}

put = EuropeanOption(
  type="put", underlying=underlying, strike=strike, dividendYield=dividendYield,
  riskFreeRate=riskFreeRate, maturity=maturity, volatility=volatility
)

put$value 
```

## Part II - Options Hedging

#### Using the options data from file `OptionsPrices.csv` used in the workshop calculate the P&L breakdown into `delta`, `gamma` and `vega` components of 1425 December 2007 put for each day between 2/27/2007 and 3/5/2007.
#### Compare actual P&L for 1425 December 2007 put with the estimated daily P&L for each day between 2/27/2007 and 3/5/2007.
#### Put the results in a table with columns: actual P&L, estimated P&L, delta P&L, gamma P&L, vega P&L.

```{r}
suppressWarnings(library(RQuantLib))

optionsData = read.csv('OptionsPrices.csv', header=T, row.names=1)
dim(optionsData) 
head(optionsData)
```


Calculate implied volatility of put with strike 1425 between 2/26/2007 and 3/5/2007

```{r}
put1425ImpliedVolatility = as.vector(
  apply(
    optionsData[1:6,], 1, function(z) EuropeanOptionImpliedVolatility(
      "put", value=z[26], underlying=z[3], strike=1425, dividendYield=z[2],
      riskFreeRate=z[2], maturity=z[1] / 365, volatility=0.3
    )
  )
) * 100

cbind(row.names(optionsData[1:6,]), put1425ImpliedVolatility)
```


Find option premium and Greeks of 1425 put between 2/26/2007 and 3/5/2007

```{r}
putValues = matrix(NA, nrow=6, ncol=4)

for(i in 1:6) {
  currentPutValue = EuropeanOption(
    type="put", underlying=optionsData[i, 3], strike=1425,
    maturity=optionsData[i, 1] / 365, riskFreeRate=optionsData[i, 2],
    volatility=put1425ImpliedVolatility[i] / 100,
    dividendYield=optionsData[i, 2]
  )
  
  putValues[i, 1] = as.numeric(currentPutValue[1])
  putValues[i, 2] = as.numeric(currentPutValue[2])
  putValues[i, 3] = as.numeric(currentPutValue[3])
  putValues[i, 4] = as.numeric(currentPutValue[4])
}

colnames(putValues) = c('value', 'delta', 'gamma', 'vega')
rownames(putValues) = row.names(optionsData[1:6, ])
putValues
```

Calculate the estimated daily P&L for each day between 2/27/2007 and 3/5/2007

```{r}
PNLDaily = matrix(NA, nrow=5, ncol=5)

for(i in 1:5) {
  # actual P&L
  PNLDaily[i, 1] = optionsData[i+1, 26] - optionsData[i, 26]
  
  # delta P&L
  PNLDaily[i, 3] = putValues[i,2] * (optionsData[i+1, 3] - optionsData[i, 3])
  
  # gamma P&L
  PNLDaily[i, 4] = 0.5 * putValues[i, 3] * (optionsData[i+1, 3] - optionsData[i, 3])^2
  
  # vega P&L
  PNLDaily[i, 5] = putValues[i, 4] / 100 * (put1425ImpliedVolatility[i+1] - put1425ImpliedVolatility[i])
  
  # estimated P&L
  PNLDaily[i, 2] = PNLDaily[i, 3] + PNLDaily[i, 4] + PNLDaily[i, 5]
}

colnames(PNLDaily) = c('actual P&L', 'estimated P&L', 'delta P&L', 'gamma P&L', 'vega P&L')
rownames(PNLDaily) = row.names(optionsData[2:6, ])

PNLDaily
```
